---
title: "Week 1 Exercises"
author: "Asher Meyers"
date: "June 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##
MARR 2.1

```{r}

playbill <- read.csv(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/playbill.csv", sep = ",")
View(playbill)
plot(playbill$LastWeek, playbill$CurrentWeek)

PlaybillFit <- lm(CurrentWeek ~ LastWeek, data = playbill)
summary(PlaybillFit)

#a) and b) 95% confidence interval for slope and intercept
confint(PlaybillFit, level = 0.95) 

#c) predict fit of $400k last week, and 95% CI
predict(PlaybillFit, data.frame(LastWeek = 400000), interval = "predict")
```

a) Yes, 1 is a plausible value for B1, because 1 is in the 95% CI

b) We fail to reject the null, because $10k is in the 95% CI for the slope.

c) $450k is not reasonable. Outside the 95% CI.

d) No, because play audience volumes decay over time - they get smaller. A better rule would be that audiences will be slightly smaller.
Two qualifiers:
1. A play that was sold out may still have enough demand that exceeds the supply, so the audience volume won't decline.
2. External events like weather and holidays may affect attendance, upsetting week on week trends.



MARR 2.2

```{r}
indicators <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/indicators.txt", header = TRUE)
View(indicators)

plot(indicators$LoanPaymentsOverdue, indicators$PriceChange)
#Find the intercept and slope of Y = price change, x = percentage of mortgage loans overdue
indicatorsFit <- lm(PriceChange ~ LoanPaymentsOverdue, data = indicators)

summary(indicatorsFit)

#a) 95% confidence interval for slope and intercept
confint(indicatorsFit, level = 0.95) 

#b) 95% CI for X = 4
predict(indicatorsFit, data.frame(LoanPaymentsOverdue = 4), interval = "predict")

```

a) 95% CI for slope: -4.163454, -0.3335853

Yes, there is evidence of a significant negative linear association, as the 95% CI is entirely negative.

b) 0% is a plausible value, because 0% is within the 95% confidence interval



MARR 2.3

```{r}
invoices <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/invoices.txt", header = TRUE)
View(invoices)

plot(invoices$Invoices, invoices$Time)
#Find the intercept and slope of Y = price change, x = percentage of mortgage loans overdue
invoicesFit <- lm(Time ~ Invoices, data = invoices)

summary(invoicesFit)

#a) and b) 95% confidence interval for slope and intercept
confint(invoicesFit, level = 0.95) 

#c) 95% CI for X = 4
predict(invoicesFit, data.frame(Invoices = 130), interval = "predict")

```


a) 95% CI for start-up time, i.e. the intercept:
0.391249620 0.89217014

b) 95% cI for B1: (0.0096, 0.013)
Fail to reject the null, as the null is in the 95% CI for the slope

c) 2.11 hours P.E., interval : (1.42, 2.80)


##LMR – 1.1, 1.3, 1.4, 1.5, 2.4, 2.5, 3.6, 4.5, 5.2


LMR 1.1

```{r}
teengamb <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/teengamb.txt", header = TRUE)
View(teengamb)

#info on dataset: http://www.life.illinois.edu/dietze/Lectures2012/Lesson22.5_Classical_regression_demo.r
# 0 = male

```

1)> plot(teengambm$verbal, teengambm$gamble)
> plot(teengambf$verbal, teengambf$gamble)
> plot(teengambm$verbal, teengambm$gamble)
> plot(teengambf$income, teengambf$gamble)
> plot(teengambm$income, teengambm$gamble)
> plot(teengambf$income, teengambf$gamble)
> plot(teengambm$income, teengambm$gamble)
> plot(teengambm$status, teengambm$gamble)
> plot(teengambf$status, teengambf$gamble)
> plot(teengambm$status, teengambm$gamble)



LMR 1.3, 2.4, 2.5


```{r}
prostate <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/prostate.txt", header = TRUE)
View(prostate)

summary(prostate)
prostateFit <- lm(lpsa ~ lcavol, data = prostate)
summary(prostateFit)



prostateFit <- lm(lpsa ~ lcavol, data = prostate) #Linear regression
rsqs <- summary(prostateFit)$r.squared #Rsquared
res <- summary(prostateFit)$sigma #Residual standard error of regression

prostateFit <- lm(lpsa ~ lcavol + lweight, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi + lbph, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi + lbph + age, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

prostateFit <- lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45 + gleason, data = prostate)
rsqs <- append(rsqs, summary(prostateFit)$r.squared)
res <- append(res, summary(prostateFit)$sigma)

rsqs <- round(rsqs, 3)
res <- round(res, 3)

plot(c(1:8), rsqs, main = "R squared of models with X number of variables", ymin = 0, ylim = c(0,1))
plot(c(1:8), res, main = "Residual standard error with X number of variables", ymin = 0, ylim = c(0,1))


pairs(~lpsa + lcavol, data = prostate)

lpsaFit <- lm(lpsa ~ lcavol, data = prostate)

lcavolFit <- lm(lcavol ~ lpsa, data = prostate)

par(cex=.8)
plot(prostate$lpsa, prostate$lcavol, xlim = c(2.45,2.5), ylim = c(1.34,1.36))
abline(lcavolFit)
abline(a = -lpsaFit$coefficients[1]/lpsaFit$coefficients[2], b = 1/lpsaFit$coefficients[2]) #Reverse the regression line to plot on same graph.
#The lines intersect at lpsa = 2.48, lcavol = 1.35


par(cex = .8)
plot(prostate$lcavol, prostate$lpsa)
abline(lpsaFit)


plot(lpsaFit)
```

LMR 1.4


```{r}
sat <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/sat.txt", header = TRUE, sep = "\t")
View(sat)

summary(sat)
plot(sat$expend, sat$total)
plot(sat$ratio, sat$total)
plot(sat$salary, sat$total)
plot(sat$takers, sat$total)
```


LMR 1.5
```{r}
divusa <- read.table(file = "C:/Users/asher/Documents/Classes/CUNY/DATA 621/Week 1/divusa.txt", header = TRUE, sep = "\t")
View(divusa)

summary(divusa)
plot(divusa$year, divusa$divorce)
plot(divusa$unemployed, divusa$divorce)
plot(divusa$femlab, divusa$divorce)
plot(divusa$marriage, divusa$divorce)
plot(divusa$birth, divusa$divorce)
plot(divusa$military, divusa$divorce)
```

LMR 3.6 - happy
```{r}
install.packages('faraway') #Install package faraway
require(faraway)
data(happy) #run in console
View(happy)

pairs(~ happy + money + sex + love + work, data = happy) #Scatter plot
happyFit <- lm(happy ~ money + sex + love + work, data = happy)
summary(happyFit)
#Only love was significant at the 1% level.
table(happy$happy)
#results are not normally distributed,

##Part C - test significance of money using permutation method

lmod <- lm(happy ~ sex + money, data = happy)
lms <- summary(lmod)
tstats <- numeric(4000)
set.seed(123)
for (i in 1:4000) {
  lmods <- lm(happy ~ sex + sample(money), data = happy)
  tstats[i] <- summary(lmods)$coef[3, 3]
}
mean(abs(tstats) > abs(lms$coef[3]))
```


LMR 4.5

```{r}
require(faraway)
data(fat)

fullFat <- lm(brozek ~ age + weight + height + neck + chest + abdom + hip +thigh + knee + ankle + biceps + forearm + wrist, data = fat)
summary(fullFat)

lowFat <- lm(brozek ~ age + weight + height + abdom, fat)
summary(lowFat)
#The observed adjusted r-squared is nearly the same, 71.7% vs 73.5%

x <- model.matrix(fullFat)
(x0 <- apply(x,2,median))
(y0 <- sum(x0*coef(fullFat)))
predict(fullFat, new = data.frame(t(x0)))
predict(fullFat, new = data.frame(t(x0)), interval = "prediction")
predict(fullFat, new = data.frame(t(x0)), interval = "confidence")

#For full model, the prediction interval is (9.62, 25.37)

xlow <- model.matrix(lowFat)
(x1 <- apply(xlow,2,median))
(y1 <- sum(x1*coef(lowFat)))
predict(lowFat, new = data.frame(t(x1)))
predict(lowFat, new = data.frame(t(x1)), interval = "prediction")
predict(lowFat, new = data.frame(t(x1)), interval = "confidence")

#The reduced model prediction interval is (9.70, 25.98) - so it's not too much bigger.


#Part c - examine cases 25-50
View(fat[25:50, ])
lowFatData <- fat[25:50, c(1, 4, 5, 6, 11)]
(residuals2550 <- abs(fat$brozek[25:50] - predict(lowFat, new = data.frame(xlow))[25:50])) #residuals = observed - predicted
residuals2550[order(-residuals2550)]

#observations 39 and 32 are the most anomalous


lowFat <- lm(brozek ~ age + weight + height + abdom, fat[c(-32,-39),])
summary(lowFat)
xlow <- model.matrix(lowFat)
(x1 <- apply(xlow,2,median))
(y1 <- sum(x1*coef(lowFat)))
predict(lowFat, new = data.frame(t(x1)))
predict(lowFat, new = data.frame(t(x1)), interval = "prediction")
predict(lowFat, new = data.frame(t(x1)), interval = "confidence")

#new interval: 9.95, 25.88 - smaller, but not a lot smaller.


(residuals <- abs(fat$brozek - predict(lowFat, new = data.frame(xlow))))
residuals[order(-residuals)][1:10]

fat[39,]
x0
```








LMR 2.6
```{r}
#Install package faraway
require(faraway)
cheddar <- data(cheddar)

pairs(~ taste + Acetic + H2S + Lactic, data = cheddar) #Scatter plot
